---
title: "data_prep"
author: "pmcdevitt"
date: "5 août 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)


rm(list=ls())

library(MASS)
library(car)
library(moments)
library(ggplot2)

home_dir <- "~/_smu/_src/blr_prj3/"
setwd(home_dir)
data_dir <- "./data"

```


```{r read data, include = TRUE, message = FALSE}

	setwd(home_dir)
	setwd(data_dir)

	shots <- read.csv("data.csv", stringsAsFactors = FALSE)
	setwd(home_dir)

	names(shots) <- tolower(names(shots))
	
	for (i in 2:(length(shots)))
	{
		if (class(shots[,i]) == "character")
		{
			shots[,i] <- factor (shots[,i])
		}
	}
```

***  

####	Some Wikipedia BBall trivia :
> In the National Basketball Association (NBA), the court is 94 by 50 feet  
The NBA adopted the three-point line at the start of the 1979–80 season. This is of variable distance, ranging from 22 feet (6.7 m) in the corners to 23.75 feet (7.24 m) behind the top of the key.  
Kobe Bean Bryant is an American retired professional basketball player and businessman. He played his entire 20-year career with the Los Angeles Lakers of the National Basketball Association (NBA). He entered the NBA directly from high school and won five NBA championships with the Lakers.  
Playing career 	1996–2016  
Career statistics  
Points 	33,643 (25.0 ppg)  
Rebounds 	7,047 (5.2 rpg)  
Assists 	6,306 (4.7 apg)  

***  

```{r remove outliers, include = TRUE, message = FALSE}

# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	remove outliers ... more than 5 sigma from mean value
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
	
	lst <- length(shots)	#
	
	for (i in 1 : lst)
	{
		if(class(shots[,i]) == "integer" || class(shots[,i]) == "numeric")
		{
			shots[,i][which(scale(shots[,i]) > 5)] <- NA
			shots[,i][which(scale(shots[,i]) < -5)] <- NA
		}
	}

```


```{r raw data tabular summary, message = FALSE}

	summary_tbl <- data.frame(x = character(0), stats = character(0))

	for (i in 2 : length(shots))
	{
		if(class(shots[,i]) == "integer" || class(shots[,i]) == "numeric")
		{
			new_row <- data.frame(x = names(shots[i]),
								  stats = sprintf (
								  	 "| %8d | %8d | %8.1f | %8.1f | %8.1f | %8.1f | %8.3f | ", 
									colSums(!is.na(shots[i])),
									(dim(shots)[1] - colSums(!is.na(shots[i]))),
								  	mean(shots[,i], na.rm = TRUE),
								  	median(shots[,i], na.rm = TRUE),
								  	max(shots[,i], na.rm = TRUE),
								  	min(shots[,i], na.rm = TRUE),
								  	skewness(shots[,i], na.rm = TRUE)
								  	)
								  )
			summary_tbl <- rbind(summary_tbl, new_row)
		}
	}

	summary_tbl
	
```


```{r add some variables}

# ...	seconds_remaining is seconds remaining in the current minute
# ...	minutes_remaining is minutes remaining in the period
# ...	there are 4 (regular time) periods in a match, each of 12 minutes
# ...	create new vairable of time remaining in the match
# ...		- creates negative values for overtime periods

	shots$time_remaining <- shots$seconds_remaining +
							shots$minutes_remaining * 60 +
							(4 - shots$period) * 12 * 60

# ...	home / away games can be determined from 'matchup' field
# ...		@ - designates away game
# ...		vs. - designates home game

	shots$home_away <- "home"
	away_lst <- grep("@", shots$matchup, perl=TRUE, value=FALSE)
	shots$home_away[away_lst] <- "away"

# ...	shot distances appear to be in feet * 10 ... just convert to feet
	
	shots$x_ft <- shots$loc_x / 10
	shots$y_ft <- shots$loc_y / 10
	
# ...	add polar coordinate - from basket to shot point
	
	shots$rad <- shots$shot_distance
	shots$ang <- atan2(shots$x_ft, shots$y_ft) * 180/pi 
	
# ...	calculate total points scored in this data set
	
	shots$pts_scored <- 2
	three_pt_lst <- grep("3PT", shots$shot_type, perl=TRUE, value=FALSE)
	shots$pts_scored[three_pt_lst] <- 3
	shots$pts_scored <- shots$pts_scored * shots$shot_made_flag
	
	total_pts_scored <- sum(shots$pts_scored, na.rm = TRUE)

```

```{r drop no information columns} 

# ...	team_id is a constant value for all rows
# ...	team_name is always Los Angeles Lakers

	shots <- subset(shots, select = -c(team_id, team_name))

```

```{r first plots}

	plot(x_ft ~ y_ft, shots, col = "blue")
	plot(x_ft ~ y_ft, shots, col = shot_distance)
	plot(x_ft ~ y_ft, shots, col = shot_zone_area)
	plot(x_ft ~ y_ft, shots, col = shot_zone_range)
	
	plot(x_ft ~ y_ft, shots, col = shot_made_flag+2)
	
	plot(x_ft ~ y_ft, data = shots[shots$shot_made_flag == 1,],
		 	col = "darkgreen",
		 	xlim = c(-10, 60))
	plot(x_ft ~ y_ft, data = shots[shots$shot_made_flag == 0,],
		 	col = "grey",
		 xlim = c(-10, 60))

	plot(x_ft ~ y_ft, shots, col = shot_type)

	plot(rad ~ ang, data = shots, col = shot_made_flag+2)
	
	boxplot(shot_distance ~ shot_made_flag, data = shots)
	boxplot(shot_distance ~ shot_type, data = shots)
	boxplot(shot_distance ~ team_id, data = shots)

```